Hypertension and microbiome

### 2019-01-04 articletwo.Rmd
Friday

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)
library(Maaslin)
library(parallel)
library(officer)
library(flextable)
library(rvg)
library(tableone)
library(scales)
library(gridExtra)
```

@@

writeimage <- function(doc, filename, header, footer, width = 7, height = 7) {
    doc <- doc %>% 
        body_add_fpar(fpar(ftext("Figure . ", prop = text.bold), ftext(header, prop = text.normal)), style = "Image Caption") %>%
        body_add_img(src = filename, width = width, height = height) %>% 
        body_add_par(footer, style = "footnote text") %>%
        body_add_break(pos = "after") 
}

@@

ggsave(file = "cache/alpha-beta.png", plot=g4, height=fig.height, width=11)
writeimage(doc, "cache/alpha-beta.png", "Alpha-beta", "footer", width = 11, height = fig.height)
```

@@

writeimage(doc, "cache/pca.png", "Alpha-beta", "footer", width = 11, height = fig.height)
```

@@

```{r}
``

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)
library(Maaslin)
library(parallel)
library(officer)
library(flextable)
library(rvg)
library(tableone)
library(scales)
library(gridExtra)
library(png)
```

@@

writeimage <- function(doc, filename, header, footer, width = 7.5) {
    img <- readPNG("cache/pca.png")
    doc <- doc %>% 
        body_add_fpar(fpar(ftext("Figure . ", prop = text.bold), ftext(header, prop = text.normal)), style = "Image Caption") %>%
        body_add_img(src = filename, width = width, height = width*dim(img)[1]/dim(img)[2]) %>% 
        body_add_par(footer, style = "footnote text") %>%
        body_add_break(pos = "after") 
}

@@

ggsave(file = "cache/alpha-beta.png", plot=g4, height=fig.height, width=11)
writeimage(doc, "cache/alpha-beta.png", "Alpha-beta", "footer")
```

@@

writeimage(doc, "cache/pca.png", "Alpha-beta", "footer", width = 7)
```

@@

```{r PCA heatmap}
pca.axis <- pca.prcomp$x %>% data.frame
pca.dset <- base::merge(meta(pseq.species), pca.axis[,1:5], by=0, all=TRUE) %>%
  dplyr::rename(Sample_ID = Row.names)

pca.dset %>% head

pca.ret <- lapply(paste0("PC", seq(1,5)), function (pc) {
    calculateglm(pca.dset,
                 modelstring = paste("%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  +", pc),
                 filterstr = "PC") %>%
        select(response, estimate, p.value) %>%
        mutate(model_name = pc)
})

pca.ret  %>%
    data.table::rbindlist() %>%
    as.data.frame %>%
    mutate(p.value = p.adjust(p.value, method="BH")) %>%
    arrange(p.value) %>%
    kable
```

```{r PCA loadings}

pca.rotation.abs <- abs(pca.prcomp$rotation)
pca.loading <- sweep(pca.rotation.abs, 2, colSums(pca.rotation.abs), "/") %>%
    as.data.frame %>%
    tibble::rownames_to_column(var = "term")

pca.pc3 <- pca.loading %>%
    select(term, PC3) %>%
    arrange(-PC3) %>%
    head(20)

pca.loading %>% pull(PC3) %>% head(2700) %>% sum

ggplot(data=pca.pc3, aes(x=reorder(term, -PC3), y=PC3)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

dev.off()

### 2019-01-04 rrbiome.Rmd
For fig 4
Friday

We calculate primary component analysis using centered log-ratio
transformation for the abunances.

@@

```{r maaslin plot, message = FALSE, echo = FALSE}
source("edgeR_plot_functions.R")

@@

```{r lactobacillus scatterplot, eval = FALSE, include = FALSE}
maaslin.lacto.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "DIAB")

prepare.maaslin(pseq.lacto,
                 tsvfile = "maaslin_lacto_merged_phfinrisk_genus.tsv",
                conffile = "maaslin_lacto_config.config",
                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.lacto.force.vars),
                readpclrows = "Lactobacillus_acetotolerans.Bacteria-")

Maaslin("maaslin_lacto_merged_phfinrisk_genus.tsv",
        "maaslin_lacto",
        strInputConfig = "maaslin_lacto_config.config",
        strModelSelection="none",
        dSignificanceLevel = 1.0, 
        fAllvAll = TRUE,
        strForcedPredictors = maaslin.lacto.force.vars)
```

>>>>

We calculate principal component analysis using centered log-ratio
transformation for the abunances.

@@

```{r maaslin plot, message = FALSE, echo = FALSE}
source("aaro/edgeR_plot_functions.R")

@@

```{r lactobacillus scatterplot, eval = FALSE, include = FALSE}
maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")

prepare.maaslin(pseq.lacto,
                 tsvfile = "maaslin_lacto_merged_phfinrisk_genus.tsv",
                conffile = "maaslin_lacto_config.config",
                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars),
                readpclrows = "Lactobacillus_delbrueckii.Bacteria,Lactobacillus_ruminis.Bacteria,Lactobacillus_acidophilus.Bacteria,Lactobacillus_fermentum.Bacteria,Lactobacillus_paracasei.Bacteria,Lactobacillus_salivarius.Bacteria")

Maaslin("maaslin_lacto_merged_phfinrisk_genus.tsv",
        "maaslin_lacto",
        strInputConfig = "maaslin_lacto_config.config",
        strModelSelection="none",
        dSignificanceLevel = 1.0, 
        fAllvAll = TRUE,
        strForcedPredictors = maaslin.force.vars)
```
